---
title: "Summer School Lausanne 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


All the scripts of the UseR sessions will be available here.

More details on the Pricing Game will follow.

## Required Software
- R (at least 3.4.3 https://cloud.r-project.org/bin/windows/base/ )
- RStudio ( https://www.rstudio.com/products/rstudio/download/#download )
- For Windows, Rtools : https://cran.r-project.org/bin/windows/Rtools/Rtools35.exe
- For Mac: Install gfortran-6.1.pkg (more info https://cran.r-project.org/bin/macosx/tools/ )



Please run the following script in your R session to install all the required packages. Make sure Rtools (Windows) or gfortan (Mac) has been installed previously.


Start with the package containing datasets. If there is a query to install from source, answer "y" (yes).
```{r,include=TRUE, message=FALSE, warning=FALSE}
if (!require("xts")) install.packages("xts")
if (!require("CASdatasets")) install.packages("CASdatasets", repos = "http://cas.uqam.ca/pub/R/", type="source")
require("CASdatasets")
```

If the installation succeeded, the following lines should load a dataset and show the first six lines.
```{r}
data("freMTPLfreq")
head(freMTPLfreq)
```

Then, we can install and load all the other packages.

```{r,include=TRUE, message=FALSE, warning=FALSE}
if (!require("Rcpp")) install.packages("Rcpp")
if (!require("zoo")) install.packages("zoo")
if (!require("sp")) install.packages("sp")
if (!require("caret")) install.packages("caret")
if (!require("mgcv")) install.packages("mgcv")
if (!require("plyr")) install.packages("plyr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("visreg")) install.packages("visreg")
if (!require("MASS")) install.packages("MASS")
if (!require("plotrix")) install.packages("plotrix")
if (!require("lme4")) install.packages("lme4")
if (!require("glmnet")) install.packages("glmnet")
if (!require("parallel")) install.packages("parallel")
if (!require("devtools")) install.packages("devtools")
if (!require("rpart")) install.packages("rpart")
if (!require("rpart.plot")) install.packages("rpart.plot")
if (!require("randomForest")) install.packages("randomForest")
if (!require("xgboost")) install.packages("xgboost")

require("Rcpp")
require("mgcv")
require("caret")
require("gridExtra")
require("plyr")
require("visreg")
require("MASS")
require("plotrix")
require("lme4")
require("glmnet")
require("parallel")
require("devtools")
require("rpart")
require("rpart.plot")
require("xgboost")

install_github("gbm-developers/gbm3")
install_github("fpechon/rfCountData")
require("gbm3")
require("rfCountData")
```



