---
title: "Brief Introduction to R and Descriptive Analysis of the Dataset"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true    
    theme: united
    highlight: tango
    code_folding: show
    keep_md: false
  github_document:
        toc: true
---
```{r, include=FALSE}
options(encoding = 'UTF-8')
#Loading all the necessary packages
if (!require("CASdatasets")) install.packages("CASdatasets")
if (!require("caret")) install.packages("caret")
if (!require("mgcv")) install.packages("mgcv")
if (!require("plyr")) install.packages("plyr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("visreg")) install.packages("visreg")
if (!require("MASS")) install.packages("MASS")
if (!require("plotrix")) install.packages("plotrix")


require("CASdatasets")
require("mgcv")
require("caret")
require("gridExtra")
require("plyr")
require("visreg")
require("MASS")
require("plotrix")


```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 75)
```

# Introduction
```{r, tidy=TRUE}
## Loading the dataset
require("CASdatasets")
data("freMTPLfreq")

freMTPLfreq = subset(freMTPLfreq, Exposure<=1 & Exposure >= 0 & CarAge<=25)

set.seed(85)
require("caret")
folds = createDataPartition(freMTPLfreq$ClaimNb, 0.5)
dataset = freMTPLfreq[folds[[1]], ]
save(dataset, file="dataset.RData")
```

A good idea is to check whether the dataset has been loaded correctly. To do this, the following tools can be used:

 - *head* allows to visualize the first 6 lines of the dataset.
```{r, tidy=TRUE}
head(dataset)
```
 - *str* allows to see the format of the different variables. We will typically distinguish numerical variables (real numbers or integers) and factors (categorical data).
```{r, tidy=TRUE}
str(dataset)
```
 - *summary* allows to compute for each variable some summary statistics.
```{r, tidy=TRUE}
summary(dataset)
```

If one needs some *help* on a function, typing a question mark and the name of the function in the console opens the help file of the function. For instance,
```{r, eval=FALSE, tidy=FALSE}
?head
```


# Descriptive Analysis of the portfolio
We will now have a descriptive analysis of the portfolio. The different variables available are *`r names(dataset)`*.

## PolicyID
The variable *PolicyID* related to a unique identifier of the policy. We can check that every policy appears only once in the dataset
```{r, tidy=TRUE}
length(unique(dataset$PolicyID)) == nrow(dataset)
```

## Exposure in month
The Exposure reveals the fraction of the year during which the policyholder is in the portfolio. 
We can compute the total exposure by summing the policyholders' exposures. Here we find `r format(sum(dataset$Exposure))` years.

We can show the number of months of exposure on a table.
```{r, fig.align='center', tidy=TRUE}
table(cut(dataset$Exposure, breaks = seq(from = 0, to = 1,by = 1/12), labels = 1:12))
```
Using the function *prop.table*, it is possible to represent this information in relative terms
show the number of months of exposure on a table.
```{r, fig.align='center', tidy=TRUE}
round(prop.table(table(cut(dataset$Exposure, breaks = seq(from = 0, to = 1,by = 1/12), labels = 1:12))), 4)
```

Alternatively, we can use a barplot !
```{r,fig.align='center', cache=TRUE, tidy=TRUE, dpi=500}
Exposure.summary = cut(dataset$Exposure, breaks = seq(from = 0, to = 1,by = 1/12))
levels(Exposure.summary) = 1:12
ggplot()+geom_bar(aes(x=Exposure.summary)) + xlab("Number of months") + ggtitle("Exposure in months")
```

## Number of claim : ClaimNb
```{r,fig.align='center', cache=TRUE, tidy=TRUE, dpi=500}
ggplot(dataset, aes(x=ClaimNb))+geom_bar()+
geom_text(stat='count', aes(label=..count..), vjust=-1)+ylim(c(0,210000))+
  ylab("")+ xlab("Number of Claims")+  ggtitle("Proportion of policies by number of claims")
```

We can compute the average claim frequency in this portfolio, taking into account the different exposures.
```{r, results='asis',eval=FALSE, tidy=TRUE}
sum(dataset$ClaimNb) / sum(dataset$Exposure)
```
Here, we obtain **`r round(sum(dataset$ClaimNb) / sum(dataset$Exposure), 4)`**.

Let us now look at the other variables.

## Power

The variable *Power* is a categorized variable, related to the power of the car. The levels of the variable are ordered categorically.
We can see the different **levels** of a **factor** by using the function *level* in R:
```{r, echo=TRUE, cache=TRUE, tidy=TRUE}
levels(dataset$Power)
```

We can see the number of observations in each level of the variable, by using the function *table*.
```{r, tidy=TRUE}
table(dataset$Power)
```
Remember however, that in insurance, exposures may differ from one policyholder to another. Hence, the table above, does NOT measure the exposure in each level of the variable *Power*. We can use the function *ddply* to give us the exposure in each level of the variable.

```{r, tidy=TRUE}
require(plyr)
Power.summary = ddply(dataset, .(Power), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure))
```

We can show this on a plot as well:
```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(Power.summary, aes(x=Power, y=totalExposure, fill=Power)) + 
  geom_bar(stat="identity")+
  ylab("Exposure in years")+
  geom_text(stat='identity', aes(label=round(totalExposure, 0), color=Power), vjust=-0.5)+
  guides(fill=FALSE, color=FALSE)
```


Let us now look at the observed claim frequency in each level
```{r, tidy=TRUE, results='hide', message=FALSE}
Power.summary = ddply(dataset, .(Power), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Frequency = sum(ClaimNb)/sum(Exposure))
Power.summary
```
```{r, results='asis', warning=FALSE, echo=FALSE, message=FALSE}
require(xtable)
print(xtable(Power.summary, digits=c(0,1,2,0,0,5)), type = "html", include.rownames=FALSE)
```


We can compute the ratio to the portfolio claim frequency.
```{r, fig.align='center', dpi=500, tidy=TRUE}
portfolio.cf = sum(dataset$ClaimNb)/ sum(dataset$Exposure)

ggplot(Power.summary) + geom_bar(stat="identity", aes(x=Power, y=Obs.Claim.Frequency, fill=Power)) + 
  geom_line(aes(x = as.numeric(Power),y=portfolio.cf), color="red") + guides(fill=FALSE)
```


## CarAge

The vehicle age, in years. This is the first continuous variable that we encounter (although it only takes discrete values).

```{r, fig.align='center', cache=TRUE, tidy=TRUE, dpi=500}
ggplot(dataset, aes(x=CarAge)) + geom_bar()  + xlab("Age of the Car")
```
Again, here, the exposures are not considered on the histogram. We can use *ddply* to correct this.
```{r, tidy=TRUE, results='hide'}
CarAge.summary = ddply(dataset, .(CarAge), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure))
CarAge.summary
```
```{r, results='asis', warning=FALSE, echo=FALSE}
require(xtable)
print(xtable(CarAge.summary,  digits=c(0,0,2,0)), type = "html", include.rownames=FALSE)
```
Then, we can plot the data onto a barplot, as before.
```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(CarAge.summary, aes(x=CarAge, y=totalExposure)) + geom_bar(stat='identity') + ylab("Exposure in years")
```

We can see a large difference, specially for new cars, which makes sense ! Indeed, let us look at the Exposure for new vehicles, using a boxplot for instance.
```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(dataset[dataset$CarAge==0,], aes(x="Exposure", y=Exposure)) + geom_boxplot() +ggtitle("Exposure of new cars")
```


Let us now also compute the claim frequency by age of car,
```{r, tidy=TRUE, results='hide'}
CarAge.summary = ddply(dataset, .(CarAge), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Freq = sum(ClaimNb)/sum(Exposure))
CarAge.summary
```
```{r, results='asis', warning=FALSE, echo=FALSE}
require(xtable)
print(xtable(CarAge.summary,  digits=c(0,1,2,0,0,5)), type = "html",include.rownames=FALSE)
```
and plot it!

```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(CarAge.summary, aes(x=CarAge, y=Obs.Claim.Freq)) + geom_point() + ylab("Observed Claim Frequency")+xlab("Age of the Car") +ylim(c(0,0.08))

```

## DriverAge
Similarly to the Age of the Car, we can visualize the Age of the Drivers.
```{r, fig.align='center', cache=TRUE, tidy=TRUE, dpi=500, results='hide'}
DriverAge.summary = ddply(dataset, .(DriverAge), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Freq = sum(ClaimNb)/sum(Exposure))
head(DriverAge.summary,9)
```

```{r, results='asis', warning=FALSE,echo=FALSE}
require(xtable)
print(xtable(head(DriverAge.summary,9), digits=c(0,1,2,0,0,5)), type = "html",include.rownames=FALSE)
```


We can show the Exposures by Age of the Driver
```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(DriverAge.summary, aes(x=DriverAge, y=totalExposure)) + geom_bar(stat='identity', width=0.8) + ylab("Exposure in years")+xlab("Age of the Driver")
```

and the observed claim frequency by Age.
```{r, fig.align="center", dpi=500, tidy=TRUE}
ggplot(DriverAge.summary, aes(x=DriverAge, y=Obs.Claim.Freq)) + geom_point()+ylab("Observed Claim Frequency") + xlab("Age of the Driver")
```

## Brand
The variable *Brand* is a categorized variable, related to the brand of the car. 
We can see the different *levels* of a *factor* by using the function **level** in R:
```{r, tidy=TRUE}
levels(dataset$Brand)
```


```{r, echo=TRUE, tidy=TRUE, results='hide'}
Brand.summary = ddply(dataset, .(Brand), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Freq = sum(ClaimNb)/sum(Exposure))
Brand.summary
```

```{r, results='asis', warning=FALSE, echo=FALSE}
require(xtable)
print(xtable(Brand.summary,  digits=c(0,1,2,0,0,5)), type = "html", include.rownames=FALSE)
```

```{r, fig.align='center', dpi=500, tidy=TRUE}
require(ggplot2)
ggplot(Brand.summary, aes(x=reorder(Brand,totalExposure), y=totalExposure, fill=Brand)) +
  geom_bar(stat='identity') +
  coord_flip()+guides(fill=FALSE)+xlab("")+ylab("Exposure in years")
```

Let us now look at the claim frequency by Brand of the car.
```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(Brand.summary, aes(x=reorder(Brand,Obs.Claim.Freq), y=Obs.Claim.Freq, fill=Brand)) +
  geom_bar(stat='identity') +
  coord_flip()+guides(fill=FALSE)+ ggtitle("Observed Claim Frequencies by Brand of the car")+xlab("")+ylab("Observed Claim Frequency")
```

## Gas
The variable *Gas* is a categorized variable, related to the fuel of the car. 
We can see the different *levels* of a *factor* by using the function **level** in R:
```{r, echo=TRUE, cache=TRUE, tidy=TRUE}
levels(dataset$Gas)
```

```{r, echo=TRUE, fig.align='center', dpi=500, tidy=TRUE}
Gas.summary = ddply(dataset, .(Gas), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Freq = sum(ClaimNb)/sum(Exposure))
ggplot(Gas.summary, aes(x=Gas, y=totalExposure, fill=Gas)) + geom_bar(stat="identity") + guides(fill=FALSE)
```

There seems to be a similar amount of Diesel and Regular gas vehicles in the portfolio.
It is generally expected that Diesel have a higher claim frequency. Does this also hold on our dataset ?
```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(Gas.summary, aes(x=Gas, y=Obs.Claim.Freq, fill=Gas)) + geom_bar(stat="identity") + guides(fill=FALSE)
```

## Region
The variable *Region* is a categorized variable, related to the region of the place of residence. 
We can see the different *levels* of a *factor* by using the function **level** in R:
```{r, echo=TRUE, cache=TRUE, tidy=TRUE}
levels(dataset$Region)
```

What are the Exposures in each region ? What are the observed claim frequencies ?

```{r, tidy=TRUE, results='hide'}
Region.summary = ddply(dataset, .(Region), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Freq = sum(ClaimNb)/sum(Exposure))
Region.summary
```

```{r, results='asis', echo=FALSE}
print(xtable(Region.summary, digits=c(0,1,2,0,0,5)), type="html", include.rownames=FALSE)
```


Using the function *twoord.plot* we can easily show both the Exposures and the observed claim frequencies on the same plot.

```{r, fig.align='center', dpi=500, tidy=TRUE}
twoord.plot(1:10,Region.summary$totalExposure,1:10,Region.summary$Obs.Claim.Freq,xlab="Region",
rylim=c(0,0.1),type=c("bar","p"), xticklab = Region.summary$Region, ylab = "Exposure", rylab = "Observed Claim Frequency")
```

We can plot a map with the observed claim frequencies

```{r, message=FALSE, fig.align='center', warning=FALSE, dpi=500, tidy=TRUE}
library(maptools)
library(ggplot2)
area <- readShapePoly("shapefiles/FRA_adm2.shp") # From http://www.diva-gis.org/gData

Region.summary$id = sapply(Region.summary$Region, substr, 2, 3)
area.points = fortify(area, region="ID_2")

area.points = merge(area.points, Region.summary[,c("id","totalExposure","Obs.Claim.Freq")], by.x = "id", by.y = "id", all.x=TRUE)
area.points = area.points[order(area.points$order),]



ggplot(area.points, aes(long, lat, group=group)) + ggtitle("Observed Claim Frequencies")+
  geom_polygon(aes(fill = area.points$Obs.Claim.Freq))+
  scale_fill_gradient(low = "yellow", high = "red", name="Obs. Claim Freq.", limits= c(0.061,0.085))+
xlab("Longitude") + ylab("Latitude")
```

and the exposures (on a log-scale)...
```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(area.points, aes(long, lat, group=group)) + ggtitle("log Exposures in years")+
  geom_polygon(aes(fill = log(area.points$totalExposure)))+
  scale_fill_gradient(low = "blue", high = "red", name="log Exposure")+
xlab("Longitude") + ylab("Latitude")
```



## Density
The Density represents here the density of the population at the place of residence.
Let us take a look at the densities in the dataset.
```{r, fig.align='center', dpi=500, tidy=TRUE}
summary(dataset$Density)
ggplot(dataset, aes(Density)) + geom_histogram(bins=200)
```

Here, contrary to the age of the driver, or the age of the car, the density has lots of different values
```{r, eval=FALSE, tidy=TRUE}
length(unique(dataset$Density))
```

We can compute this by using the command above, and we get `r length(unique(dataset$Density))`.
```{r, tidy=TRUE, results='hide'}
Density.summary = ddply(dataset, .(Density), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Freq = sum(ClaimNb)/sum(Exposure))
head(Density.summary)
```

```{r, results='asis', echo=FALSE, fig.align='center'}
print(xtable(head(Density.summary), digits=c(0,1,2,0,0,5)), type="html", include.rownames=FALSE)
```

We can plot the observed claim frequencies...

```{r, fig.align='center', dpi=500, tidy=TRUE}
ggplot(Density.summary, aes(x=Density, y=Obs.Claim.Freq)) + geom_point()
```

... but realize it is impossible to see a trend. One way out is to categorize the variable. We will see later (GAM) that it is possible to estimate a smooth function, which avoid the arbitrary categorization.

We can categorize the variable using the function *cut*.
```{r, tidy=TRUE}
dataset$DensityCAT = cut(dataset$Density, breaks = quantile(dataset$Density, probs = seq(from = 0, to = 1, by=0.1)),
                         include.lowest = TRUE)
table(dataset$DensityCAT)
levels(dataset$DensityCAT) <- LETTERS[1:10]
```
Then, we can apply the same strategy as above.


```{r, tidy=TRUE, results='hide'}
Density.summary = ddply(dataset, .(DensityCAT), summarize, totalExposure = sum(Exposure), Number.Observations = length(Exposure), Number.Claims = sum(ClaimNb), Obs.Claim.Freq = sum(ClaimNb)/sum(Exposure))
Density.summary
```

```{r, results='asis', echo=FALSE}
print(xtable(Density.summary, digits=c(0,1,2,0,0,5)), type="html", include.rownames=FALSE)
```


Using the function *twoord.plot* we can easily show both the Exposures and the observed claim frequencies on the same plot.


```{r, fig.align='center', dpi=500, tidy=TRUE}
twoord.plot(1:10,Density.summary$totalExposure,1:10,Density.summary$Obs.Claim.Freq,xlab="Density (categorized)", lylim=c(0,15000),
rylim=c(0,0.15),type=c("bar","p"), xticklab = Density.summary$Density, ylab = "Exposure", rylab = "Observed Claim Frequency", lytickpos=seq(0,15000,5000), rytickpos=seq(0,0.15,0.03))
```
